A Survey of Physical Design Techniques of Information Systems

### 2010-07-08 trunk/cs848-db/Paper.tex

Partitioning is an important aspect of physical design as it reduces table scanning time. There are two main categories of partitioning: horizontal partitioning and vertical partitioning. Horizontal partitioning divides the table into sets of table rows where each row or record still has all the attributes of the table. For example, dividing the data by date where all data dating less than year 2000 lies in one partition while all data dating more lies in another. On the other hand, vertical partitioning reduces the width of the dataset by storing some attributes in one table, and some in another.

The main advantage of any type of partitioning is to reduce the amount of time it takes to scan a table which in turn improves performance. For example, if a table has 1000 records, and a query is only using the first 20 records then having these 20 records in a partition containing 50 records will save time as opposed to examining all the 1000 records.

\subsection{Clustering}
When only records satisfying certain criteria are need from a table, it is often unnecessary to scan the whole table. For example, if only records that occrurred in January 2010 are needed, then querying all the table would be an extremely unnecessary overhead. This is where clustering comes in. Clustering relfects how the records are physically located together on disk. Records that are likely to be queried together will be stored physically together to avoid querying the whole table. For the previously mentioned example, the data can be clustered by month, such that all the records of January 2010 are physically in the same page, for example. In this case, only this page will need to be read. As people realized the potential of clustering, multidimensional clustering was introduced where data is clustered based on different criteria at the same time.

@@

@@

@@

- joins in data warehouses are usualy expensive since the fact table (the largest one) participates in every join \cite{628286}

@@

\subsubsection{Materlized Views}
not very useful since most queries are ad-hoc \cite{653447}
\cite{355309}

@@

hierachical data clusters \cite{karayannidis2008hierarchical}

>>>>

Partitioning is an important aspect of physical design as it reduces table scanning time. There are two main categories of partitioning: horizontal partitioning
and vertical partitioning. Horizontal partitioning divides the table into sets of table rows where each row or record still has all the attributes of the table.
For example, dividing the data by date where all data dating less than year 2000 lies in one partition while all data dating more lies in another. On the other
hand, vertical partitioning reduces the width of the dataset by storing some attributes in one table, and some in another.
The main advantage of any type of partitioning is to reduce the amount of time it takes to scan a table which in turn improves performance. For example, if a
table has 1000 records, and a query is only using the first 20 records then having these 20 records in a partition containing 50 records will save time as
opposed to examining all the 1000 records. Further classification in partitioning is single vs group horizontal or vertical
partitioning~\cite{gruenwald1990database}. Horizontal partitioning just divides the data into partitions where each group contains complete tuples. Group
horizontal partitioning groups tuples that are more frequently used together. Single vertical partitioning divides a relation into groups each of which contains
only one type of attribute. Group vertical partitioning divides the data vertically such that attributes that are commonly required together are physically
stored together.

\subsection{Clustering}
When only records satisfying certain criteria are need from a table, it is often unnecessary to scan the whole table. For example, if only records that occurred
in January 2010 are needed, then querying all the table would be an extremely unnecessary overhead. This is where clustering comes in. Clustering reflects how
the records are physically located together on disk. Records that are likely to be queried together will be stored physically together to avoid querying the
whole table. For the previously mentioned example, the data can be clustered by month, such that all the records of January 2010 are physically in the same
page, for example. In this case, only this page will need to be read. As
people realized the potential of clustering, multidimensional clustering was introduced where data is clustered based on different criteria at the same time.

@@

cite~\cite{dewitt1984implementation}.

@@

Partitioning is used in disk based systems to divide the data into the physical pages in a way that will yield the best performance. This could be through
vertical partitioning or horizontal partitioning, as previously explained. In main memory databases, we do not have the issue of expensive disk access which
makes partitioning in memory unnecessary as it will not save any costs. However, partionining is needed in the secondary storage of the data on
disk~\cite{gruenwald1990database}. Disk storage is still used as a backup of main memory databases. When a crash occurs, a reload of the database from disk or
archive memory takes place. In order to avoid page faults caused by referencing data that is still being reloaded, the reloading process must be designed to be
efficient in loading the important data such that the database users are not affected. Gruenwald and Eich~\cite{gruenwald1990database,gruenwald1990choosing}

@@

- joins in data warehouses are usually expensive since the fact table (the largest one) participates in every join \cite{628286}

@@

\subsubsection{Materialized Views}
not very useful since most queries are ad-hoc \cite{653447}
\cite{355309}

@@

hierarchical data clusters \cite{karayannidis2008hierarchical}
